Closing the gap between statistical and scientific workflows for improved forecasts in ecology

### 2025-03-27 forecastflows.tex
some writing

>>>>

\vspace{0.2cm}
\noindent \textbf{Scientific practice and current workflows}

% General scientific method we all learn
Quantitative science rely on a model-based framework to confront hypothesis and data, making some approximations \citep{}. The general scientific method stresses out that the research question should guide both the design of the experiment and the corresponding model building. The experiment should then be conducted---according to this specific design. Finally, the resulting experimental data should be used to inform our model and differentiate between hypotheses---hopefully answering our initial question. However, this is often an idealized view and does not reflect the complexity of ecology \citep{}.

% Divergences from this are common 
Divergences from this scientific method are common. One explanation is the lack of rigor, leading to questionable practices, such as retrospectively crafting hypothesis to explain the results of a model rather than testing a clear pre-defined question (data-driven analysis, Fig. 1). But the reality of ecological research is also a major driver of the divergences from the ideal scientific method. Many important questions often cannot be addressed by conducting experiments and replications, and we must often rely on existing datasets to have a large-scale and long-term perspective. However, this does not explain the persistent flaws and lack of overall coherence in ecological modeling. Trend estimation still often relies on a one-way model fitting, and most of the time is spent fitting the model to empirical data (Fig. 1). For forecasting, researchers focus on making predictions with complex models, but the steps of model building and parameterization are not very transparent and not clearly delineated (Fig. 1). The different parts of the model are often calibrated separately rather than as a whole, and some parameter values are just fixed based on experiments and expert knowledge.

% There is thus room for improvment, with a workflow that addresses theses flaws while taking account the reality of working with ecological data.
